\documentclass[11pt,dvipdfm]{article}

\usepackage[utf8]{inputenc}
\usepackage{deauthor,times,graphicx,hyperref}
\usepackage{subfig}
\usepackage{paralist}
\usepackage{multirow}
\usepackage{macros}

%for highlight changes
\usepackage{soul}

\begin{document}

\title{Geo-Replication: Fast If Possible, Consistent If Necessary\thanks{Student author are followed by faculty author names, both in alphabetical order. Cheng Li and Valter Balegas are the lead authors of the work.}}

\author{Valter Balegas$^1$, Cheng Li$^2$, Mahsa Najafzadeh$^4$, Daniel Porto$^{3}$, 
Allen Clement$^{2,5}$, Sérgio Duarte$^1$, \\
Carla Ferreira$^1$, Johannes Gehrke$^6$, Jo\~{a}o Leit\~{a}o$^1$, Nuno Preguiça$^1$, Rodrigo Rodrigues$^3$, \\
Marc Shapiro$^4$, Viktor Vafeiadis$^2$\\
\small $^1$NOVA LINCS, DI, FCT, Universidade NOVA de Lisboa
$^2$Max Planck Institute for Software Systems (MPI-SWS)\\
\small $^3$INESC-ID / IST, University of Lisbon 
$^4$Inria \& Sorbonne Universités, UPMC Univ Paris 06, LIP6\\
\small $^5$Currently at Google
$^6$Microsoft and Cornell University}

\graphicspath{{./figures/redblue/redblueOrder/}{./figures/}}

\maketitle

\input{abstract}
\input{intro}
\input{model}
\input{consistency}
\input{convergence}
\input{invariant}
\input{related}
\input{conclude}

\begin{thebibliography}{10} 
\itemsep=1pt 
\begin{small}


\bibitem{dc-outages}
{7 Data Center Disasters You'll Never See Coming}.
\newblock
  \url{http://www.informationweek.com/cloud/7-data-center-disasters-youll-never-see-coming/d/d-id/1320702}.
\newblock Accessed Feb-2016.

\bibitem{riak}
Using data types -- riak documentation.
\newblock \url{http://docs.basho.com/riak/latest/dev/using/data-types/}.
\newblock Accessed Feb-2016.

\bibitem{soundcloud}
Consistency without Consensus: CRDTs in Production at SoundCloud.
\newblock  \url{http://www.slideshare.net/InfoQ/consistency-without-consensus-crdts-in-production-at-soundcloud}
\newblock Accessed Feb-2016.

\bibitem{chainreaction}
S.~Almeida, J.~Leit\~{a}o, and L.~Rodrigues.
\newblock {Chain{R}eaction: A Causal+ Consistent Datastore Based on Chain
  Replication}.
  \newblock In {\em EuroSys '13}, 85--98, 2013. ACM.
%\newblock In {\em Proc.\ 8th ACM European Conf.\ on Computer Systems}, EuroSys
%  '13, pages 85--98, New York, NY, USA, 2013. ACM.

\bibitem{ramp}
P.~Bailis, A.~Fekete, J.~M. Hellerstein, A.~Ghodsi, and I.~Stoica.
\newblock {Scalable Atomic Visibility with RAMP Transactions}.
\newblock In {\em SIGMOD '14}, 27--38, 2014. ACM.
%\newblock In {\em Proc.\ 2014 ACM SIGMOD Int.\ Conf.\ on Management of Data},
%  SIGMOD '14, pages 27--38, New York, NY, USA, 2014. ACM.

\bibitem{bolton}
P.~Bailis, A.~Ghodsi, J.~M. Hellerstein, and I.~Stoica.
\newblock {Bolt-on Causal Consistency}.
\newblock In {\em SIGMOD '13}, 761--772, 2013. ACM.
%  \newblock In {\em Proc.\ 2013 ACM SIGMOD Int.\ Conf.\ on Management of Data},
%  SIGMOD '13, pages 761--772, New York, NY, USA, 2013. ACM.

\bibitem{Balegas2015Indigo}
V.~Balegas, S.~Duarte, C.~Ferreira, R.~Rodrigues, N.~Pregui\c{c}a,
  M.~Najafzadeh, and M.~Shapiro.
\newblock Putting consistency back into eventual consistency.
\newblock In {\em EuroSys '15}, 6:1--6:16, 2015. ACM.
%  \newblock In {\em Proceedings of the Tenth European Conference on Computer
%  Systems}, EuroSys '15, pages 6:1--6:16, New York, NY, USA, 2015. ACM.

\bibitem{Balegas2015Counter}
V.~Balegas, D.~Serra, S.~Duarte, C.~Ferreira, M.~Shapiro, R.~Rodrigues, and
  N.~Preguica.
\newblock Extending eventually consistent cloud databases for enforcing numeric
  invariants.
\newblock In {\em SRDS '15}, 31--36, Sept 2015.
%  \newblock In {\em Reliable Distributed Systems (SRDS), 2015 IEEE 34th Symposium
%  on}, pages 31--36, Sept 2015.

\bibitem{BarbaraMilla1994Demarcation}
D.~Barbar\'{a}-Mill\'{a} and H.~Garcia-Molina.
\newblock The demarcation protocol: A technique for maintaining constraints in
  distributed database systems.
\newblock {\em The VLDB Journal}, 3(3):325--353, July 1994.

\bibitem{Bernstein1987CCR}
P.~A. Bernstein, V.~Hadzilacos, and N.~Goodman.
\newblock {\em Concurrency Control and Recovery in Database Systems}.
\newblock Addison-Wesley, 1987.

\bibitem{pnuts}
B.~F. Cooper, R.~Ramakrishnan, U.~Srivastava, A.~Silberstein, P.~Bohannon,
  H.-A. Jacobsen, N.~Puz, D.~Weaver, and R.~Yerneni.
\newblock {PNUTS: Yahoo!'s Hosted Data Serving Platform}.
\newblock {\em Proc. VLDB Endow.}, 1(2):1277--1288, Aug. 2008.

\bibitem{spanner}
J.~C. Corbett, J.~Dean, M.~Epstein, A.~Fikes, C.~Frost, J.~J. Furman,
  S.~Ghemawat, A.~Gubarev, C.~Heiser, P.~Hochschild, W.~Hsieh, S.~Kanthak,
  E.~Kogan, H.~Li, A.~Lloyd, S.~Melnik, D.~Mwaura, D.~Nagle, S.~Quinlan,
  R.~Rao, L.~Rolig, Y.~Saito, M.~Szymaniak, C.~Taylor, R.~Wang, and
  D.~Woodford.
\newblock {Spanner: Google's Globally-distributed Database}.
\newblock In {\em OSDI '12}, 251--264, 2012. USENIX
  Association.
%\newblock In {\em Proc.\ 10th USENIX Conf.\ on Operating Systems Design and
%  Implementation}, OSDI'12, pages 251--264, Berkeley, CA, USA, 2012. USENIX
%  Association.

\bibitem{dynamo}
G.~DeCandia, D.~Hastorun, M.~Jampani, G.~Kakulapati, A.~Lakshman, A.~Pilchin,
  S.~Sivasubramanian, P.~Vosshall, and W.~Vogels.
\newblock {Dynamo: Amazon's Highly Available Key-value Store}.
\newblock In {\em  SOSP '07}, 205--220, 2007. ACM.
%\newblock In {\em Proc.\ 21st ACM SIGOPS Symp.\ on Operating Systems
% Principles}, SOSP '07, pages 205--220, New York, NY, USA, 2007. ACM.

\bibitem{orbe}
J.~Du, S.~Elnikety, A.~Roy, and W.~Zwaenepoel.
\newblock {Orbe: Scalable Causal Consistency Using Dependency Matrices and
  Physical Clocks}.
\newblock In {\em  SOCC '13}, 11:1--11:14, 2013. ACM.
%  \newblock In {\em Proc.\ 4th Annual Symp.\ on Cloud Computing}, SOCC '13, pages
%  11:1--11:14, New York, NY, USA, 2013. ACM.

\bibitem{semanticDDB}
H.~Garcia-Molina.
\newblock Using semantic knowledge for transaction processing in a distributed
  database.
\newblock {\em ACM Trans. Database Syst.}, 8(2):186--213, June 1983.

\bibitem{GotsmanConsistencyReason}
A.~Gotsman, H.~Yang, C.~Ferreira, M.~Najafzadeh, and M.~Shapiro.
\newblock 'cause i'm strong enough: Reasoning about consistency choices in
  distributed systems.
\newblock In {\em POPL 2016}, 371--384, 2016. ACM.
%  \newblock In {\em Proceedings of the 43rd Annual ACM SIGPLAN-SIGACT Symposium
%  on Principles of Programming Languages}, POPL 2016, pages 371--384, New York,
%  NY, USA, 2016. ACM.

\bibitem{mdcc}
T.~Kraska, G.~Pang, M.~J. Franklin, S.~Madden, and A.~Fekete.
\newblock {MDCC: Multi-data Center Consistency}.
\newblock In {\em EuroSys '13}, 113--126, 2013. ACM.
%  \newblock In {\em Proc.\ 8th ACM European Conf.\ on Computer Systems}, EuroSys
%  '13, pages 113--126, New York, NY, USA, 2013. ACM.

\bibitem{Ladin1992LazyReplication}
R.~Ladin, B.~Liskov, L.~Shrira, and S.~Ghemawat.
\newblock Providing high availability using lazy replication.
\newblock {\em ACM Trans. Comput. Syst.}, 10(4):360--391, Nov. 1992.

\bibitem{cassandra}
A.~Lakshman and P.~Malik.
\newblock {Cassandra: A Decentralized Structured Storage System}.
\newblock In {\em SIGOPS Oper. Syst. Rev.}, 44(2):35--40, 2010.

\bibitem{Li14Automating}
C.~Li, J.~Leit{\~a}o, A.~Clement, N.~Pregui{\c c}a, R.~Rodrigues, and
  V.~Vafeiadis.
\newblock Automating the choice of consistency levels in replicated systems.
\newblock In {\em ATC '14},
  281--292, 2014. USENIX Association.
%  \newblock In {\em 2014 USENIX Annual Technical Conference (USENIX ATC 14)},
%  pages 281--292, Philadelphia, PA, June 2014. USENIX Association.

\bibitem{Li2012RedBlue}
C.~Li, D.~Porto, A.~Clement, J.~Gehrke, N.~Pregui\c{c}a, and R.~Rodrigues.
\newblock {Making Geo-replicated Systems Fast As Possible, Consistent when
  Necessary}.
\newblock In {\em OSDI '12}, 265--278, 2012. USENIX
  Association.
%\newblock In {\em Proc.\ 10th USENIX Conf.\ on Operating Systems Design and
%  Implementation}, OSDI'12, pages 265--278, Berkeley, CA, USA, 2012. USENIX
%  Association.  
  

\bibitem{warranties}
J.~Liu, T.~Magrino, O.~Arden, M.~D. George, and A.~C. Myers.
\newblock Warranties for faster strong consistency.
\newblock In { \em NSDI '14}, 2014. USENIX Association.
%\newblock In {\em Proc.\ 11th USENIX Conf.\ on Networked Systems Design and
%  Implementation}, {NSDI'14}, Berkeley, CA, USA, 2014. USENIX Association.

\bibitem{cops}
W.~Lloyd, M.~J. Freedman, M.~Kaminsky, and D.~G. Andersen.
\newblock {Don't Settle for Eventual: Scalable Causal Consistency for Wide-area
  Storage with COPS}.
\newblock In { \em SOSP  '11}
, 401--416, 2011. ACM.
%\newblock In {\em Proc.\ 23d ACM Symp.\ on Operating Systems Principles}, SOSP
%  '11, pages 401--416, New York, NY, USA, 2011. ACM.

\bibitem{eiger}
W.~Lloyd, M.~J. Freedman, M.~Kaminsky, and D.~G. Andersen.
\newblock {Stronger Semantics for Low-latency Geo-replicated Storage}.
\newblock In {\em  NSDI '13}, 313--328, 2013. USENIX
  Association.
%  \newblock In {\em Proc.\ 10th USENIX Conf.\ on Networked Systems Design and
%  Implementation}, {NSDI'13}, pages 313--328, Berkeley, CA, USA, 2013. USENIX
%  Association.

\bibitem{Mahajan2010Depot}
P.~Mahajan, S.~Setty, S.~Lee, A.~Clement, L.~Alvisi, M.~Dahlin, and M.~Walfish.
\newblock Depot: cloud storage with minimal trust.
\newblock In {\em OSDI}, 2010.

\bibitem{replicatedcommit}
H.~Mahmoud, F.~Nawab, A.~Pucher, D.~Agrawal, and A.~El~Abbadi.
\newblock {Low-latency Multi-datacenter Databases Using Replicated Commit}.
\newblock {\em Proc. VLDB Endow.}, 6(9):661--672, 2013.

\bibitem{ONeil1986Escrow}
P.~E. O'Neil.
\newblock The escrow transactional method.
\newblock {\em ACM Trans. Database Syst.}, 11(4):405--430, Dec. 1986.

\bibitem{mobisnap}
N.~Pregui\c{c}a, J.~L. Martins, M.~Cunha, and H.~Domingos.
\newblock {Reservations for Conflict Avoidance in a Mobile Database System}.
\newblock In {\em MobiSys '03}, 43--56, 2003. ACM.
%  \newblock In {\em Proc.\ 1st Int.\ Conf.\ on Mobile Systems, Applications and
%  Services}, MobiSys '03, pages 43--56, New York, NY, USA, 2003. ACM.

\bibitem{Roy14Writes}
S.~Roy, L.~Kot, G.~Bender, B.~Ding, H.~Hojjat, C.~Koch, N.~Foster, and
  J.~Gehrke.
\newblock The homeostasis protocol: Avoiding transaction coordination through
  program analysis.
\newblock In {\em SIGMOD '15}, 1311--1326, 2015.
%\newblock In {\em Proceedings of the 2015 {ACM} {SIGMOD} International
%  Conference on Management of Data, Melbourne, Victoria, Australia, May 31 -
%  June 4, 2015}, pages 1311--1326, 2015.  

\bibitem{Schurman2009latency}
E.~Schurman and J.~Brutlag.
\newblock Performance related changes and their user impact. {P}resented at
  velocity web performance and operations conference.
\newblock \url{http://slideplayer.com/slide/1402419/}, 2009.

\bibitem{crdts}
M.~Shapiro, N.~Pregui\c{c}a, C.~Baquero, and M.~Zawirski.
\newblock {Conflict-free Replicated Data Types}.
\newblock In {\em SSS '11}, 386--400, 2011. Springer-Verlag.
%\newblock In {\em Proc.\ 13th Int.\ Conf.\ on Stabilization, Safety, and
%  Security of Distributed Systems}, SSS'11, pages 386--400, Berlin, Heidelberg,
%  2011. Springer-Verlag.  
  

\bibitem{exo-leasing}
L.~Shrira, H.~Tian, and D.~Terry.
\newblock {Exo-leasing: Escrow Synchronization for Mobile Clients of Commodity
  Storage Servers}.
\newblock In {\em Middleware '08}, 42--61, 2008. Springer-Verlag New
  York, Inc.
%\newblock In {\em Proc.\ 9th ACM/IFIP/USENIX Int.\ Conf.\ on Middleware},
%  Middleware '08, pages 42--61, New York, NY, USA, 2008. Springer-Verlag New
%  York, Inc.

\bibitem{Singh2009Zeno}
A.~Singh, P.~Fonseca, P.~Kuznetsov, R.~Rodrigues, and P.~Maniatis.
\newblock {Zeno: Eventually Consistent Byzantine-Fault Tolerance.}
\newblock In {\em NSDI'09}, 169--184, 2009.
%  \newblock In {\em Proceedings of the 6th USENIX Symposium on Networked Systems
%  Design and Implementation (NSDI'09)}, pages 169--184, 2009.

\bibitem{dynamoDB}
S.~Sivasubramanian.
\newblock {Amazon DynamoDB: A Seamlessly Scalable Non-relational Database
  Service}.
\newblock In {\em SIGMOD '12}, 729--730, 2012. ACM.
%  \newblock In {\em Proc.\ 2012 ACM SIGMOD Int.\ Conf.\ on Management of Data},
%  SIGMOD '12, pages 729--730, New York, NY, USA, 2012. ACM.

\bibitem{walter}
Y.~Sovran, R.~Power, M.~K. Aguilera, and J.~Li.
\newblock {Transactional Storage for Geo-replicated Systems}.
\newblock In {\em SOSP  '11}, 385--400, 2011. ACM.
%\newblock In {\em Proc.\ 23d ACM Symp.\ on Operating Systems Principles}, SOSP
%  '11, pages 385--400, New York, NY, USA, 2011. ACM.  
  

\bibitem{Pileus}
D.~B. Terry, V.~Prabhakaran, R.~Kotla, M.~Balakrishnan, M.~K. Aguilera, and
  H.~Abu-Libdeh.
\newblock {Consistency-based Service Level Agreements for Cloud Storage}.
\newblock In {\em SOSP '13}, 309--324, 2013. ACM.
% \newblock In {\em Proc.\ 24th ACM Symp.\ on Operating Systems Principles}, SOSP
%  '13, pages 309--324, New York, NY, USA, 2013. ACM.

\bibitem{bayou}
D.~B. Terry, M.~M. Theimer, K.~Petersen, A.~J. Demers, M.~J. Spreitzer, and
  C.~H. Hauser.
\newblock {Managing Update Conflicts in Bayou, a Weakly Connected Replicated
  Storage System}.
\newblock In {\em SOSP  '95}, 172--182, 1995. ACM.
%  \newblock In {\em Proc.\ 15th ACM Symp.\ on Operating Systems Principles}, SOSP
%  '95, pages 172--182, New York, NY, USA, 1995. ACM.

\bibitem{Walborn95Semantics}
G.~D. Walborn and P.~K. Chrysanthis.
\newblock {Supporting Semantics-based Transaction Processing in Mobile Database
  Applications}.
\newblock In {\em SRDS '95}, 31--40, 1995. IEEE Computer Society.
%\newblock In {\em Proc.\ 14th Symp.\ on Reliable Distributed Systems}, SRDS
%  '95, pages 31--40, Washington, DC, USA, 1995. IEEE Computer Society.  
  

\bibitem{Zawirski15Write}
M.~Zawirski, N.~Pregui\c{c}a, S.~Duarte, A.~Bieniusa, V.~Balegas, and
  M.~Shapiro.
\newblock Write fast, read in the past: Causal consistency for client-side
  applications.
\newblock In {\em Middleware '15}, 75--87, 2015. ACM.
%\newblock In {\em Proceedings of the 16th Annual Middleware Conference},
%  Middleware '15, pages 75--87, New York, NY, USA, 2015. ACM.  
  

\bibitem{transactionchains}
Y.~Zhang, R.~Power, S.~Zhou, Y.~Sovran, M.~K. Aguilera, and J.~Li.
\newblock {Transaction Chains: Achieving Serializability with Low Latency in
  Geo-distributed Storage Systems}.
\newblock In {\em SOSP '13}, 276--291, 2013. ACM.
%  \newblock In {\em Proc.\ 24th ACM Symp.\ on Operating Systems Principles}, SOSP
%  '13, pages 276--291, New York, NY, USA, 2013. ACM.


\end{small} 
\end{thebibliography}

\end{document}

